### 2014-06-02 patchy-selection-paper.tex
edited haplotype length section; added haplotypes to figure

\begin{figure}[ht]
  \begin{center}
    \includegraphics{sim-transit.pdf}
  \end{center}
  \caption{
  BLAH BLAH WIGGLY LINE
  \plr{can we please keep that caption? ok, fine\dots}
  \textbf{Adaptation by migration:}
  An initially unadapted patch (10 demes at the left end of the range) is colonized around generation 5000;
  a representative lineage from the new patch is shown transiting between the two and coalescing with a lineage
  in the unadapted patch.
  As in Figure~\ref{fig:sim_snapshots}, color and contours show allele frequency across space and time;
  other simulation parameters are as in Figures~\ref{fig:sim_occupation_freqs}B and \ref{fig:sim_snapshots}.
  \label{fig:lineagesmotion}
  }
\end{figure}

@@

If a patch adapts through new mutation or a rare migrant lineage, the
genomic region surrounding the selected site will hitchhike along with
it \citep{maynardsmith1974hitchhiking},
so that at least initially, all adapted individuals within a patch
share a fairly long stretch of haplotype. 
Pairs of individuals within a patch will share a haplotype of mean genetic length of
$\approx \log(\rho A s_p)/s_p$ around the selected site, 
as long as the patch is reasonably well mixed by dispersal
\citep[otherwise see][]{barton2013genetic}.
This association gets slowly whittled down by recombination during interbreeding at the edge of the patch;
but there will always be longer LD nearby to the selected site.

When an already adapted patch colonizes another through migration,
then the newly colonized patch will inherit a long piece of haplotype around the selected site from the originating patch.
The genetic length of this haplotype is roughly inversely proportional to the
number of generations the allele spends ``in transit'', 
because each reproductive event in locations where the allele is at
low frequency recombines haplotypes of the $b$ allele closer to the
adaptive allele. In this case, the haplotype is literally hitchhiking across geography!
\gc{In Figure \ref{fig:lineagesmotion} we show the lineage that founds a
mutant population on a second patch. The entire time that the lineage
is outside the dark red zone (where the B allele is common) the
haplotype that accompanies it is being broken down. After the lineage
establishes on the patch the rate of decay of the haplotype is
significant slowed as it is again surrounded by other B
gneotypes with the similar haplotypic backgrounds. We first discuss
the transit of the lineage, and its accompanying haplotype, and return
to this point about the whittling down of haplotypes within
estalbished patches in the next section.}

@@

Since each generation spent in transit provides an opportunity for recombination,
if recombination is Poisson, the length of the haplotype (in Morgans)
initially shared between populations on each side of the selected locus is exponentially distributed
with mean $\tau$, so that if $L$ is the length of hitchhiking segment on, say, the right of
the selected locus, then
\begin{equation} \label{eqn:haplen_cdf2}
\P\{L>\ell\} = \E[e^{-\ell \tau}|\tau<\infty] ,
%  &= \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+s_m)} - \sqrt{2s_m}\right)}\right\} .
\end{equation}
and so (one minus) equation~\eqref{eqn:haplen_cdf} in fact gives the cumulative distribution function of $L$.
The form of this distribution function implies that if $Y$ is an exponential random variable with rate $R\sqrt{2}/\sigma$,
then $L$ has the same distribution as $(Y + \sqrt{s_m})^2 - s_m$.
Furthermore, the expected length of shared hitchiking haplotype is
\begin{equation} \label{eqn:haplotype_length}
\E[L] = \sigma \sqrt{2s_m}/R + \sigma^2/R^2
\end{equation}
and $\var[L] = 2s_m\sigma^2/R^2 + 4 \sigma^3 \sqrt{2s_m}/R^3 + 5 \sigma^4 / R^4$.
For two dimensions, asymptotics of Bessel functions show that $L$ has the same tail behavior,
but how other properties might change is unclear.

As a general rule of thumb, equation~\eqref{eqn:mean_tau} means that families who successfully establish
move at speed $\sqrt{s_m} \sigma$ towards the new patch 
-- if $s_m$, the strength of the selection against them, is smaller, the need to move quickly is less imperative.
Then, equation~\eqref{eqn:haplotype_length} means that the haplotype length
is roughly the length one would expect given the mean transit time.
Note that we have become used to seeing $\sigma$ divided by $\sqrt{s_m}$, rather than multiplied,
the $\sigma/\sqrt{s_m}$ is a length scale, and to convert it to a time this is divided by $1/s_m$.
\gc{In the Appendix Section \ref{apx:limiting_dist_of_HH_haplotype} we
  give some expressions for the limiting distribution of
  the length of shared haplotype.}

\gc{Graham pasted this to the appendix as it felt like too much information to drop on
  people}
%%%

@@

The other key issue in thinking about the population genetic signal of
adaptation to spatial patches is the process by which haplotypes are
whittled down within patches, which plays into a number of aspects of
haplotype sharing. Thee initial haplotype of the sweep within a patch will
diserpate over time as recombination brokes up the haplotype; likewise the haplotype that is shared 
between patches that have
adapted through a shared migrant allele
(eqn. \ref{eqn:haplotype_length}) will also breakdown. 
We first develop a simple approximation to the rate of
recombination controling the breakdown of haplotypes within patches. 
If we wait long enough after the initial sweep,
either due to migration or mutation, the initial pattern will have
decayed due to this process of recombination. However, within patches
we should expect to find a larger extent of haplotype sharing close to
the selected locus, than in surrounding geographic regions as
selection against migrant lineage slows the diserpation of
genetic drift accumulated within the patch \citep{bartonpaper,charlesworths}.

We refer the reader again to Figure \ref{fig:lineagesmotion}. When a lineage in
established patch is in the dark red area, where the B allele is common, it is very unlikely to
recombine off the B background. However, when it wanders out of the
patch into the paler red and white areas when it recombines it will
usually be on to the b allele background.

A detailed treatment of this is well beyond the scope of this paper,
but we can get a good idea of the genomic scale of these effects through heuristic arguments.
To understand the probability of a recombination moving our lineage
off of the B background we need to understand what fraction of the
time our lineage spends in geographic regions where $b$ in common. 
Consider the lineage of a single B allele, moving back through time,
in a patch at migration-selection balance, 
i.e.\ with the local proportion of B alleles near $x$ equal to $p(x)$,
and that the lineage stays close enough to the patch 
that fluctuations about $p(x)$ are small. (Note that this assumption differs from
previous sections, but here our focus is on a typical lineage in the
patch rather than the rare lineages that make it far from the patch).
If we suppose that type B individuals at location $y$ reproduce at rate $1+s(y)$,
and the chance their offspring move to $x$ is $m(y,x)$ (with $\sum_x m(y,x)=1$),
then the total rate of influx of B alleles from $y$ into $x$ is $\rho p(y) (1+s(y)) m(y,x)$.
As argued by \citet{hudson}, the rate at which a B lineage moves from $x$ to $y$
is this influx divided by the local abundance of $B$ alleles,
or $\frac{ p(y) (1+s(y)) m(y,x) }{ p(x) }$.
After the same large-population, continuous-space limit that underlies the rest of this paper,
this motion should converge to a diffusion process \citep[as assumed by][]{hallatschek},
notwithstanding some technical difficulties (e.g.\ discontinuity of $s(x)$).
In It\^o differential notation, 
the location $X_t$ of a B lineage $t$ generations ago satisfies
\begin{align}
    d X_t &= (1+s(X_t)) \frac{d}{d\,x} \log( p(X_t) (1+s(X_t)) ) dt + \sigma \sqrt{1+s(X_t)} dB_t,\\
    % &=  ( \log(p(X_t))' (1+s(X_t)) + s'(X_t) ) dt + \sqrt{1+s(X_t)} dB_t,\\
\end{align}
where $s(x)=s_p$ if $x$ is in the patch and $s(x)=s_m$ otherwise, 
and $B_t$ is a standard Brownian motion.
This is a diffusion describing the movement of our lineage backward in
timethat, when at $x$, 
is pulled toward regions with more B alleles
by the drift term $(1+s(x))\frac{d}{d\,x} \log( p(x) (1+s(x)) )$,
and moves at nearly constant speed (since $\sqrt{1+s(x)}$ is close to 1).
Furthermore, it is easy to verify that the stationary distribution of this process is
\begin{align}
  \pi(x) = \frac{ p(x)^2 (1+s(x)) }{ \int_y p(y)^2 (1+s(y)) dy } . \label{eqn:lineagestatdist}
\end{align}
A lineage at a locus at genetic distance $r$ that begins linked to a B allele
moves as a B lineage up until the most recent time
it recombined from a b background to a B background,
which occurs when at $x$ with rate $r (1-p(x)) (1+s(x)/2)$.
The Feynman-Kac formula could then be used to write down differential equations
solved by various moments of the haplotype length as a function of the sampling location of the lineage.

As our goal here to describe the rough genomic length scale, 
based on the above, we view 
the lineage is simply bouncing around in the patch backward in time and we propose a
simple heurstic to describe the probability of recombination. 
If recombination is on a slower time scale than mixing within the patch,
then the time until the last recombination is the integral of $r (1-p(x)) (1+s(x)/2)$ 
against the stationary distribution (eqn. \ref{eqn:lineagestatdist}).
Within the patch, $p(x) \approx 1$,
so recombination off the $B$ background is unlikely to occur,
and so the rate of recombination depends to a first approximation 
on the proportion of the patch in which $p(x)$ is intermediate,
i.e. were both $B$ and $b$ are present.
This region where $p(x)$ is intermediate has width of order $\sigma/\sqrt{s}$, where $s = \min(s_p,s_m)$.
As we increase the size of a patch, the area on which $p(x)$ is intermediate scales with the perimeter of the patch.
Therefore, the proportion of the time a lineage at stationarity is found in the intermediate regions
multiplied by $p(x)$ in these intermediate regions, where we assume
$p(x) \approx 1/2$. Under these approximations the per generation rate of
recombination of a linked lineage to the b background is, taking in this region, roughly 
% \begin{equation}
% r (\sigma/\sqrt{s})A^{-1/d} \label{eqn:rate-rec-in-patch}
% \end{equation}
\begin{align}
&=
    \begin{cases}
    r \frac{\frac{\sigma}{\sqrt{s}} }{\frac{\sigma}{\sqrt{s}} + A }    \qquad & d=1 \\
   r \frac{\pi  \frac{\sigma}{\sqrt{s}} \left( \frac{\sigma}{\sqrt{s}} +
  \sqrt{\frac{A}{\pi}} \right)}{A+\pi  \frac{\sigma}{\sqrt{s}} \left( \frac{\sigma}{\sqrt{s}} +
  \sqrt{\frac{A}{\pi}} \right) }    \qquad & d=2 \\  
\end{cases}
\label{eqn:rate-rec-in-patch}   %(\sigma/\sqrt{s})A^{-1/d}
\end{align}
We can use this to first of all think about the length of haplotype
associated with the $B$ allele after a given length of time $\tau$,
this could the the time since the sweep within a patch due to a
mutation or migration of a beneficial allele. If the recombination
rate per base is $r_{BP}$ then the length of haplotype associated with
the $B$ allele is
\begin{equation}
\frac{\sqrt{s}}{\sigma A^{-1/d}}\frac{1}{r_{BP}}
\end{equation} 
note that the length of haplotype decays at a rate proportional to the
cline width $\sqrt{s} / \sigma$, as the narrower the cline the less
opportunity for recombination to move our lineage off the $B$
background.

Finally we can consider the expected length of haplotype shared within
a patch. To do this, we compare the rate of recombination between B
and b haplotypes (eqn. \ref{eqn:rate-rec-in-patch})
to the rate of coalescence of B haplotypes within a patch. We will
assume that mixing within the patch is fast compared to the rate of
drift, so that the rate of coalescence of B alleles within a patch is 
$1/(A\rho)$ 
(this bypasses the well-known difficulties to spatial coalescence \citep{felsenstein1978pain,bartonetheridgedepaulis}).
The rate of recombination and coalescence will be on the same order if
$\xi /(A\rho) \approx r (\sigma/\sqrt{s}) A^{-1/d}$, 
and hence the genetic length of haplotype shared between pairs of
individuals within the patches is on the order of
\begin{align}
  r = \frac{\sqrt{s} \xi}{\sigma \rho A^{(d-1)/d}} 
\end{align}
which does not increase with $A$ in one dimension,
but decreases with $\sqrt{A}$ in two dimensions .

@@

\section{Limiting distribution of the transit time and hitchhiking haplotype}
\label{apx:limiting_dist_of_HH_haplotype}

In the above we already assumed that the inter-patch distance $R$ is large relative to the migration-selection cline width $\sigma/\sqrt{s_m}$;
by taking the large-$R$ limit above we can obtain a more intuitive description of the process
via a central limit theorem for $\tau$.
Define $\eta_\tau$ to be the deviation of $\tau$ from its mean, normalized by its standard deviation,
so that $\tau = R/\sigma(2s_m)^{-1/2} + \eta_\tau \sqrt{R/\sigma} (2s_m)^{-3/4}$.
Plugging this into equation~\eqref{eqn:haplen_cdf},
after some algebra (and using that $\sqrt{R+\epsilon} = \sqrt{R} + \frac{\epsilon}{2\sqrt{R}} - \frac{\epsilon^2}{2\sqrt{R}^3} + O(\epsilon^3)$) we obtain that
\begin{align}
  \E[e^{-\ell \eta}] &= \exp\left\{ \frac{\ell^2}{2} + O\left(\frac{1}{R}\right) \right\}.
\end{align}
Since the Laplace transform uniquely determines a distribution \citep{durrett1996probability},
this implies that $\eta_\tau$ converges in distribution to a standard Gaussian random variable as $R \to \infty$.
In other words, we now know that the distribution of $\tau$ about its mean is approximately Gaussian.
Since the standard deviation of $\tau$ is of lower order than its mean,
a similar calculation shows that for large $R$, the haplotype halflength $L$ is approximately exponentially distributed, 
with mean $\sqrt{2s_m} \sigma/R$.

>>>>

\begin{figure}[ht]
  \begin{center}
    \includegraphics{sim-transit.pdf}
  \end{center}
  \caption{
  BLAH BLAH WIGGLY LINE
  \plr{can we please keep that caption? ok, fine\dots}
  \textbf{Adaptation by migration:}
  An initially unadapted patch (10 demes at the left end of the range) is colonized around generation 5000;
  a representative lineage from the new patch is shown transiting between the two and coalescing with a lineage
  in the unadapted patch.
  In the left panel, as in Figure~\ref{fig:sim_snapshots}, color and contours show allele frequency across space and time;
  other simulation parameters are as in Figures~\ref{fig:sim_occupation_freqs}B and \ref{fig:sim_snapshots}.
  The right panel shows (in red) the length of the shared haplotype between the two lineages
  around the selected locus,
  which decreases as time progresses.
  \label{fig:lineagesmotion}
  }
\end{figure}

@@

If a patch adapts through new mutation or a rare migrant lineage, the
genomic region surrounding the selected site will hitchhike along with it \citep{maynardsmith1974hitchhiking},
so that at least initially, all adapted individuals within a patch
share a fairly long stretch of haplotype. 
Pairs of individuals within a patch will share a haplotype of mean genetic length of
about $\log(\rho A s_p)/s_p$ around the selected site, 
as long as the patch is reasonably well mixed by dispersal
\citep[otherwise see][]{barton2013genetic}.
This association gets slowly whittled down by recombination during interbreeding at the edge of the patch;
but there will always be longer LD nearby to the selected site.

When an already adapted patch colonizes another through migration,
then the newly colonized patch will inherit a long piece of haplotype around the selected site from the originating patch.
The genetic length of this haplotype is roughly inversely proportional to the
number of generations the allele spends ``in transit'', 
because each reproductive event in locations where the allele is at
low frequency will likely recombine haplotypes of the $b$ allele closer to the adaptive allele. 
In this case, the haplotype is literally hitchhiking across geography!
Figure~\ref{fig:lineagesmotion} shows a simulation of such an event,
including the lineage that founds an adaptive population on a second patch,
and the length of the haplotype shared between this lineage and one in the original patch.
The entire time that the lineage is outside the region where the B allele is common (dark red in Figure~\ref{fig:lineagesmotion}), 
the haplotype that accompanies it is broken down rapidly. 
After the lineage establishes on the patch, the rate of decay of the haplotype is slowed significantly, 
since most others with which it recombines have similar haplotypic backgrounds. 
We first discuss the transit of the lineage, and its accompanying haplotype, and return
to how haplotypes are whittled down within estalbished patches in the next section.

@@

Since each generation spent in transit provides an opportunity for recombination,
if recombination is Poisson, the length of the haplotype (in Morgans)
initially shared between populations on each side of the selected locus is exponentially distributed
with mean $\tau$, so that if $L$ is the length of hitchhiking segment on, say, the right of
the selected locus, then
\begin{equation} \label{eqn:haplen_cdf2}
\P\{L>\ell\} = \E[e^{-\ell \tau}|\tau<\infty] ,
%  &= \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+s_m)} - \sqrt{2s_m}\right)}\right\} .
\end{equation}
and so (one minus) equation~\eqref{eqn:haplen_cdf} in fact gives the cumulative distribution function of $L$.
The form of this distribution function implies that if $Y$ is an exponential random variable with rate $R\sqrt{2}/\sigma$,
then $L$ has the same distribution as $(Y + \sqrt{s_m})^2 - s_m$.
Furthermore, the expected length of shared hitchiking haplotype is
\begin{equation} \label{eqn:haplotype_length}
\E[L] = \sigma \sqrt{2s_m}/R + \sigma^2/R^2
\end{equation}
and $\var[L] = 2s_m\sigma^2/R^2 + 4 \sigma^3 \sqrt{2s_m}/R^3 + 5 \sigma^4 / R^4$.
For two dimensions, asymptotics of Bessel functions show that $L$ has the same tail behavior;
but how other properties might change is unclear.

As a general rule of thumb, equation~\eqref{eqn:mean_tau} means that families who successfully establish
move at speed $\sqrt{s_m} \sigma$ towards the new patch 
-- if $s_m$, the strength of the selection against them, is smaller, the need to move quickly is less imperative.
Then, equation~\eqref{eqn:haplotype_length} means that the haplotype length
is roughly the length one would expect given the mean transit time.
Note that we have become used to seeing $\sigma$ divided by $\sqrt{s_m}$, rather than multiplied;
it appears here because $\sigma/\sqrt{s_m}$ is a length scale, 
and to convert it to a time this is divided by $1/s_m$.
This intuitive picture can be expanded by a central limit theorem:
for large $R$, the deviation of $\tau$ about its mean is asymptotically Gaussian,
and the haplotype halflength $L$ is exponential.

% In Appendix~\ref{apx:limiting_dist_of_HH_haplotype}, 
% we also give some expressions for the limiting distribution of the length of shared haplotype.

\gc{Graham pasted this to the appendix as it felt like too much information to drop on people}
\plr{I am sympathetic, but it's only a paragraph\dots how about that last sentence, instad?}
%%%
% 
% \section{Limiting distribution of the transit time and hitchhiking haplotype}
% \label{apx:limiting_dist_of_HH_haplotype}
% 
% In the above we already assumed that the inter-patch distance $R$ is large relative to the migration-selection cline width $\sigma/\sqrt{s_m}$;
% also taking the large-$R$ limit, obtains a more intuitive description of the process
% via a central limit theorem for $\tau$.
% Define $\eta_\tau$ to be the deviation of $\tau$ from its mean, normalized by its standard deviation,
% so that $\tau = R/\sigma(2s_m)^{-1/2} + \eta_\tau \sqrt{R/\sigma} (2s_m)^{-3/4}$.
% Plugging this into equation~\eqref{eqn:haplen_cdf},
% after some algebra (and using that $\sqrt{R+\epsilon} = \sqrt{R} + \frac{\epsilon}{2\sqrt{R}} - \frac{\epsilon^2}{2\sqrt{R}^3} + O(\epsilon^3)$) we obtain that
% \begin{align}
%   \E[e^{-\ell \eta}] &= \exp\left\{ \frac{\ell^2}{2} + O\left(\frac{1}{R}\right) \right\}.
% \end{align}
% Since the Laplace transform uniquely determines a distribution \citep{durrett1996probability},
% this implies that $\eta_\tau$ converges in distribution to a standard Gaussian random variable as $R \to \infty$.
% In other words, we now know that the distribution of $\tau$ about its mean is approximately Gaussian.
% Since the standard deviation of $\tau$ is of lower order than its mean,
% a similar calculation shows that for large $R$, the haplotype halflength $L$ is approximately exponentially distributed, 
% with mean $\sqrt{2s_m} \sigma/R$.

@@

The other key issue in thinking about the population genetic signal of adaptation in spatial patches 
is the process by which haplotypes are whittled down within patches, 
which plays into a number of aspects of haplotype sharing. 
The initial haplotype that sweeps within a patch 
will be dispersed over time by recombination;
likewise, the haplotype that is shared between patches coadapted by migration
(equation~\ref{eqn:haplotype_length}) will also break down. 
We first develop a simple approximation to the rate 
at which recombination breaks down haplotypes within patches. 
If we wait long enough after the initial adaptation of a patch,
the haplotype that swept initially will have decayed due to this process of recombination. 
However, we still expect to find individuals within the patch sharing long haplotypes around the selected locus
than with individuals elsewhere,
since selection against migrants 
decreases mean coalescence times within the patch near the selected locus.
\citep{bartonpaper,charlesworths}.

Refer again to Figure~\ref{fig:lineagesmotion}. 
When a lineage is in a region with high frequency of B alleles (dark red), 
it is very unlikely to recombine off the B background. 
However, when it wanders out of the patch,
into regions with lower frequency of B alleles,
when it recombines it will
usually be on to the background of the b allele.

A detailed treatment of this is well beyond the scope of this paper,
but we can get a good idea of the genomic scale of these effects through heuristic arguments.
To understand the probability of a recombination moving our lineage
off of the B background we need to understand what fraction of the
time our lineage spends in geographic regions where $b$ in common. 
Consider the lineage of a single B allele, moving back through time,
in a patch at migration-selection balance, 
i.e.\ with the local proportion of B alleles near $x$ equal to $p(x)$,
and that the lineage stays close enough to the patch 
that fluctuations about $p(x)$ are small. 
(Note that this assumption differs from
previous sections, but here our focus is on a typical lineage in the
patch rather than the rare lineages that make it far from the patch).
If we suppose that type B individuals at location $y$ reproduce at rate $1+s(y)$,
and the chance their offspring move to $x$ is $m(y,x)$ (with $\sum_x m(y,x)=1$),
then the total rate of influx of B alleles from $y$ into $x$ is $\rho p(y) (1+s(y)) m(y,x)$.
As argued by \citet{hudson}, the rate at which a B lineage moves from $x$ to $y$
is this influx divided by the local abundance of $B$ alleles,
or $\frac{ p(y) (1+s(y)) m(y,x) }{ p(x) }$.
After the same large-population, continuous-space limit that underlies the rest of this paper,
this motion should converge to a diffusion process \citep[as assumed by][]{hallatschek},
notwithstanding some technical difficulties (e.g.\ discontinuity of $s(x)$).
In It\^o differential notation, 
the location $X_t$ of a B lineage $t$ generations ago satisfies
\begin{align}
    d X_t &= (1+s(X_t)) \frac{d}{d\,x} \log( p(X_t) (1+s(X_t)) ) dt + \sigma \sqrt{1+s(X_t)} dB_t,\\
    % &=  ( \log(p(X_t))' (1+s(X_t)) + s'(X_t) ) dt + \sqrt{1+s(X_t)} dB_t,\\
\end{align}
where $s(x)=s_p$ if $x$ is in the patch and $s(x)=s_m$ otherwise, 
and $B_t$ is a standard Brownian motion.
This is a diffusion describing the movement of our lineage backward in
timethat, when at $x$, 
is pulled toward regions with more B alleles
by the drift term $(1+s(x))\frac{d}{d\,x} \log( p(x) (1+s(x)) )$,
and moves at nearly constant speed (since $\sqrt{1+s(x)}$ is close to 1).
Furthermore, it is easy to verify that the stationary distribution of this process is
\begin{align}
  \pi(x) = \frac{ p(x)^2 (1+s(x)) }{ \int_y p(y)^2 (1+s(y)) dy } . \label{eqn:lineagestatdist}
\end{align}
A lineage at a locus at genetic distance $r$ that begins linked to a B allele
moves as a B lineage up until the most recent time
it recombined from a b background to a B background,
which occurs when at $x$ with rate $r (1-p(x)) (1+s(x)/2)$.
The Feynman-Kac formula could then be used to write down differential equations
solved by various moments of the haplotype length as a function of the sampling location of the lineage;
instead, we only use the above as a guide for rough approximation.

Based on this picture of
the lineage bouncing around in the patch,
we propose a simple heurstic to describe the probability of recombination. 
If recombination is on a slower time scale than mixing within the patch,
then the time until the last recombination is the integral of $r (1-p(x)) (1+s(x)/2)$ 
against the stationary distribution (equation~\ref{eqn:lineagestatdist}).
Within the patch, $p(x) \approx 1$,
so recombination off the $B$ background is unlikely to occur,
and so the rate of recombination depends to a first approximation 
on the proportion of the patch in which $p(x)$ is intermediate,
i.e.\ where both $B$ and $b$ are present.
This region where $p(x)$ is intermediate has width of order $\sigma/\sqrt{s}$, where $s = \min(s_p,s_m)$.
As we increase the size of a patch, the area on which $p(x)$ is intermediate scales with the perimeter of the patch.
Therefore, the proportion of the time a lineage at stationarity is found in the intermediate regions
multiplied by $p(x)$ in these intermediate regions, where we assume
$p(x) \approx 1/2$. Under these approximations the per generation rate of
recombination of a linked lineage to the b background is
reduced by a factor roughly equal to the proportion of the patch lying within distance $\sigma/\sqrt{s}$ of the boundary,
i.e.
% \begin{equation}
% r (\sigma/\sqrt{s})A^{-1/d} \label{eqn:rate-rec-in-patch}
% \end{equation}
\plr{Note to ourselves: If a $d$-dimensional object with area A has radius $r$ (in any sense)
and you shrink it to have radius $r_0$,
then the resulting object has area $A \times (r_0/r)^d$;
here we have $r_0 = r - \sigma/\sqrt{s}$,
in $d=1$ the radius is $A/2$ and in $d=2$ it is $\sqrt{A/\pi}$.}
\begin{align}
  r_\text{eff} &=
    \begin{cases}
      r \frac{2\sigma}{A\sqrt{s}}    \qquad & d=1 \\
      r \frac{\sqrt{\pi} \sigma }{ \sqrt{ s A } } \qquad & d=2 \\  
\end{cases} ,
\label{eqn:rate-rec-in-patch}   %(\sigma/\sqrt{s})A^{-1/d}
\end{align}
i.e.\ the reduction in recombination rate is proportional to $(\sigma/\sqrt{s})A^{-1/d}$.
We can use this to think about the length of the haplotype
associated with the $B$ allele after a given length of time $T$,
for instance, the time since the $B$ allele joined the patch.
\plr{changed this to genetic lengths so as not to introduce another parameter, $r_{BP}$}
% If the recombination rate per base is $r_{BP}$, 
% then the length of the haplotype associated with the $B$ allele is
% \begin{equation}
% \frac{\sqrt{s}}{\sigma A^{-1/d}}\frac{1}{r_{BP}}
% \end{equation} 
In units of genetic length (Morgans),
the length of the haplotype associated with the $B$ allele is, up to a constant,
\begin{equation}
  \frac{A^{1/d}\sqrt{s}}{\sigma T}
\end{equation} 
Note that the length of haplotype decays at a rate proportional to the
cline width $\sqrt{s} / \sigma$, as the narrower the cline the less
opportunity for recombination to move our lineage off the $B$ background.

Finally we can consider the expected length of haplotype shared within a patch. 
To do this, we compare the rate of recombination between B and b haplotypes (equation~\ref{eqn:rate-rec-in-patch})
to the rate of coalescence of B haplotypes within a patch. 
We will assume that mixing within the patch is fast compared to the rate of drift, 
so that the rate of coalescence of B alleles within a patch is 
$\xi^2/(A\rho)$,
the ``variance effective population size'' of the patch
(this bypasses the well-known difficulties to spatial coalescence \citep{felsenstein1978pain,bartonetheridgedepaulis}).
The rate of recombination and coalescence will be on the same order if
$\xi^2 /(A\rho) \approx r (\sigma/\sqrt{s}) A^{-1/d}$, 
and hence the genetic length of haplotype shared between pairs of
individuals within the patches is on the order of
\begin{align}
  r = \frac{\sqrt{s} \xi^2}{\sigma \rho A^{(d-1)/d}} 
\end{align}
which does not depend on $A$ in one dimension,
but decreases with $\sqrt{A}$ in two dimensions.

@@
